








%!TEX root = ./main.tex

\section{Evaluating the discovery}
\label{sec:evaluation}

% \btext{@all? (in particolare chiara DF, FAbrizio, Marco e sergio per pensarla). Pensare alla valutazione. Ho recuperato i due dataset dei danesi (vedi git). Se volete si possono usare questi, pero' uno dei due log in realta' e' un insieme di 215 logs di cui loro hanno fatto la media. Forse io ne sceglierei due o tre (es Dreyers Foundation.xes + uno o due dei 215) e farei la valutazione su questi. Anche perche' piu' che valutazione e' far vedere cosa esce perche' nn mi e' chiaro come valutiamo n termini di metriche l'output.}

In Section \ref{sec:example}, we introduced the preference types through simple toy-like examples. The interested reader, however, might wonder about the usability and efficacy of our approach when applied to real-life cases. We explored the applicability of our approach using %to
two real-life event logs, namely DREYERS (492 positive traces and 208 negative ones) and CERV (55 positive traces and 102 negatives traces).\footnote{For reproducibility purposes the source code is available in~\cite{zenodo:experiments}, the DREYERS event log can be found in~\cite{DBLP:conf/ssci/DeboisS15,DBLP:conf/bpm/SlaatsDB21}, while the CERV event log is a proprietary dataset.} In both cases, we were able to find ten models satisfying the given preferences in a computation time between 1 and 3 seconds, using a normally-equipped laptop.



\subsection{The DREYERS log}
\label{subsec:dreyers}

The DREYERS log describes the Dreyer Foundation’s processes pertaining to their support to legal and architectural projects, and it has been used in \cite{DBLP:conf/ssci/DeboisS15,DBLP:conf/bpm/SlaatsDB21}. Each application to request the Foundation's support goes through a pre-screen that can lead to an initial rejection. The remaining applications undergo a review, in which at least one of the reviewers must be a lawyer or an architect, depending on the application type. The review phase is followed by a board meeting, where applications to be supported are selected and eventually funded.
Two sets of log traces are available in the dataset: a positive one collecting executions that did not fail and a negative one representing executions that were reset due to a system failure. %process instances that execute properly have been labeled as the positive example set, while process instances that do not adhere to the proper execution path belong to the negative example set. %The labeling was already available, and we are not aware of the adopted criteria.
%Two sets of log traces are available: process instances that do not go through the pre-screen belong to the negative example set, whereas the ones that execute till the end of the procedure have been labeled as positive.

Using this dataset, we played a sort of ``investigation game'', and explored the hypothesis that the type of application (architect- or lawyer- type) might affect the process outcome.
%More generally, one might be interested in knowing if there are any differences in models that explicitly mention one of the two possible types of application.
%To this end, we first specified a preference for models containing activity \textsf{Lawyer Review}, and then a preference for models containing activity \textsf{Architect Review}.
\federico{
To this end, we initially specified a preference \texttt{good\_action(Lawyer Review)}, and later on a preference \texttt{good\_action(Architect Review)}.
}
%
In both cases, more than one model %several models
satisfying the preferences were found. However, the two sets of models are identical (except for the architect/lawyer activity), showing that the process is independent of the application domain.
We report an example of a model obtained when specifying the preference for models containing activity \textsf{Architect Review}:
%
%l'ordine di restituzione è molto diverso e i primi restituiti sono quelli con l'azione in danese, quindi ne ho scelti due volutamente diversi... è meglio metterli uguali con solo l'azione architect/lawyer diversa?

{\small{
\begin{align*}
%\mathsf{M_1} = \{ \ & \mathsf{ alternateresponse(Undo\ payment, First\ payout)} \\
M_1 = \{ \ & \mathsf{ alternateresponse(Undo\ payment, First\ payout)} \\
& \mathsf{ chainprecedence(Fill\ out\ application, Initial\ Rejection)} \\
& \mathsf{ choice(Round\ ends, Change\ phase\ to\ Abort)} \\
& \mathsf{ notchainsuccession(Receive\ final\ report, First\ payout)} \\
& \mathsf{ notchainsuccession(Change\ phase\ to\ Preparation, Approve\ application)} \\
& \mathsf{ notchainsuccession(Change\ phase\ to\ Preparation, Execute\ Pre\ decision)} \\
& \mathsf{ notchainsuccession(Set\ to\ Pre\ approved, Round\ Ends)} \\
& \mathsf{ notsuccession(Architect\ Review, Approval\ on\ to\ the\ board)} \ \} 
\end{align*}
}}

%%%The DREYERS log consists of 492 positive traces and 208 negative ones, 130 positive paths and 29 negative ones. NegDis is able to return models that satisfy the whole positive log and violate all but three negative traces (two paths). 
%%%For the models of the previous example we report the number of traces and paths violated by each constraint:

%%%Model M1:\\
\begin{table}[t!]
\label{table:resultsModelM1}

\begin{center}
\begin{scriptsize}
\scalebox{1}{
\begin{tabular}{lcc}
\toprule
%\hline
\textbf{Constraints} & \textbf{Traces \#} & \textbf{Variants \#} \\
\hline
alternateresponse(Undo payment, First payout) & 2 & 2\\
%\hline
chainprecedence(Fill out application, Initial Rejection) & 3 & 2\\
%\hline
choice(Round ends, Change phase to Abort) & 195 & 17 \\
%\hline
notchainsuccession(Receive final report, First payout) & 1 & 1 \\
%\hline
notchainsuccession(Change phase to Preparation, Approve application) & 1 & 1\\
%\hline
notchainsuccession(Change phase to Preparation, Execute Pre decision) & 2 & 2 \\
%\hline
notchainsuccession(Set to Pre approved, Round Ends) & 2 & 2\\
%\hline
notsuccession(Architect Review, Approval on to the board) & 1 & 1\\
\hline
Traces not ruled out by the model & 3 & 2 \\
\hline
Total & 208 & 30\\
\hline
\end{tabular}
}
\end{scriptsize}
\end{center}
\caption{Traces ruled out by each constraint of model $M_1$.}%$\mathsf{M_1}$.}
\end{table}
%
Notably, as shown in Table 1, this model is able to discriminate between positive and negative examples except for three negative traces (two variants), that cannot be ruled out without discarding also some positive examples.

We continued our investigation by focusing on %the very beginning of the process, and on the specific 
activity \textsf{Initial Rejection}. We report here one of the returned models: %Many models were found, among them the following one:

{\small{\begin{align*}
%\mathsf{M_2} = \{ \ &  \mathsf{ absence2(Initial\ rejection)} \\
M_2 = \{ \ &  \mathsf{ absence2(Initial\ rejection)} \\
& \mathsf{ choice(Round\ Ends, Applicant\ informed)} \\
& \mathsf{ notchainsuccession(Set\ to\ Pre\ approved, Round\ Ends)} \\
& \mathsf{ notchainsuccession(Receive\ final\ report, First\ payout)} \\
& \mathsf{ notchainsuccession(Change\ phase\ to\ Preparation, Approve\ application)} \\
& \mathsf{ notchainsuccession(Change\ phase\ to\ Preparation, Execute\ Pre\ decision)} \\
& \mathsf{ notsuccession(Lawyer\ Review, Change\ phase\ to\ review)} \\
& \mathsf{ response(Undo\ payment, First\ payout) } \ \} 
\end{align*}
}
}
Model $M_2$ %$\mathsf{M_2}$
highlights the fact that some negative traces can be distinguishable from the positive ones because of the repetition of \textsf{Initial Rejection}: some traces, indeed, reported the execution of the activity twice, thus indicating an attention point for the process analyst.

\chiara{Finally, we did compare the effect of discovering models with or without the two preferred activities. For this we asked \nd to extract 10 optimal
%\todo{Lasciamo optimal?}
models with no preferences, 10 with the  \textsf{Architect Review} preference and 10 with the \textsf{Initial Rejection} preference, and we pairwise compared the models with no preference and the ones with a preferred activity. When imposing no preferences, activity \textsf{Architect Review} shows up in only 4 of the 10 discovered models. Imposing the usage of \textsf{Initial Rejection} is instead ``unnecessary'' (a posteriori), as this activity is also present in all 10 models discovered without specifying any preference.}

%%%Model M2:\\
%%%
%%%\begin{tabular}{|l|l|l|}
%%%\hline
%%%Constraint & Traces & Paths \\
%%%\hline
%%%absence2(Initial rejection) & 3 & 2\\
%%%\hline
%%%choice(Round ends, Change phase to Abort) & 195 & 17 \\
%%%\hline
%%%notchainsuccession(Receive final report, First payout) & 1 & 1 \\
%%%\hline
%%%notchainsuccession(Change phase to Preparation, Approve application) & 1 & 1\\
%%%\hline
%%%notchainsuccession(Change phase to Preparation, Execute Pre decision) & 2 & 2 \\
%%%\hline
%%%notchainsuccession(Set to Pre approved, Round Ends) & 2 & 2\\
%%%\hline
%%%notsuccession(Lawyer Review, Change phase to review) & 1 & 1\\
%%%\hline
%%%response(Undo payment, First payout) & 2 & 2\\
%%%\hline
%%%\end{tabular}
%%
%%
%%
%%%\btext{DA FINIRE DOPO AVER PARLATO CON GIULIA}



\subsection{Evaluation on the CERV log}
\label{subsec:cerv}

CERV is an event log that describes a process pertaining to the cervical cancer screening in an Italian screening center, and it has been used in previous works \cite{2007b-Lamma,deviant-tkde}. The screening program is composed of five phases, organized sequentially: screening planning, invitation management, first level test with pap-test, second level test with colposcopy (only if the first test is positive), and eventually biopsy (if the second test gives a positive response). Several subjects do not respect the planned protocol: e.g., subjects might not show up at the first test, even if they have chosen a time slot. Moreover, a number of subjects prefer to consult physicians they trust more, in case of a positive response. As it commonly happens in socio-technical systems, a large variety of process instances appear in the log, not all them being compliant with the protocol.
%
Hence, the traces have been labeled by a domain expert as belonging either to the positive or the negative set, depending on their compliance with the adopted protocol.

We investigated the log by eliciting two preferences over the \textsf{precedence} and \textsf{succession} templates:
\federico{
\begin{center}
\texttt{good\_constraint(precedence).}\\
\texttt{good\_constraint(succession).}
\end{center}
}
%.
The first two returned models are:

\begin{align*}
%\mathsf{M_1} = \{ \ &  \mathsf{ alternateresponse(send\ positive\ pap\ test\ result, take\ a\ colposcopy\ examination)} \\
M_1 = \{ \ &  \mathsf{ alternateresponse(send\ positive\ pap\ test\ result, take\ a\ colposcopy\ examination)} \\
& \mathsf{ chainprecedence(invite, take\ a\ pap\ test\ examination)} \\
& \mathsf{ exclusivechoice(send\ pap\ test\ sample, reject)} \\
& \mathsf{precedence(send\ colposcopy\ uncertain\ result, send\ biopsy\ sample)} \ \}  \\ \\
%\mathsf{M_2} = \{ \ &  \mathsf{ alternateresponse(send\ positive\ pap\ test\ result, take\ a\ colposcopy\ examination)} \\
M_2 = \{ \ &  \mathsf{ alternateresponse(send\ positive\ pap\ test\ result, take\ a\ colposcopy\ examination)} \\
& \mathsf{ chainprecedence(invite, take\ a\ pap\ test\ examination)} \\
& \mathsf{ exclusivechoice(send\ pap\ test\ sample, reject)} \\
& \mathsf{succession(send\ colposcopy\ uncertain\ result, send\ biopsy\ sample)} \ \}  
\end{align*}

In model $M_1$, the \textsf{precedence} constraint implies that if a biopsy is executed, then the colposcopy examination has provided an uncertain result before. The second model is identical to the first one, except for the constraint related %subjected
to our preference. Interestingly, the \textsf{succession} relates %is about
the same activities involved in %by
the \textsf{precedence} constraint in the first model. The difference between the two models lies in the logical relation between \textsf{precedence} and \textsf{succession}: a trace that violates the former will always violate the latter (but not vice versa% the opposite
). It is then up to the domain expert to prefer a stricter or a more general behavior.
%, knowing that in this case such a choice would not affect the resulting model.

\chiara{Finally, we did compare the effect of discovering models with or without the two preferred templates, by extracting 10 optimal
%\todo{Lasciamo optimal?}
models with no template preferences, and 10 each with the \textsf{Precedence} and \textsf{Succession} preference respectively. Interestingly enough, imposing the \textsf{Precedence} preference results in being extremely useful in this scenario. In fact, none of the 10 models discovered with no preferred template did contain a \textsf{Precedence} pattern. Similarly with \textsf{Succession}, which appears in only 1 of the 10 models discovered without specifying any preference.}

%%%For the models of the previous example we report the number of traces and paths violated by each constraint:
%%%
%%%Model M1:\\
%%%
%%%\begin{tabular}{|l|l|l|}
%%%\hline
%%%Constraint & Traces & Paths \\
%%%\hline
%%%alternateresponse(send positive pap test result, take a colposcopy examination) & 4 & 1\\
%%%\hline
%%%chainprecedence(invite, take a pap test examination) & 3 & 2\\
%%%\hline
%%%exclusivechoice(send pap test sample, reject) & 94 & 2 \\
%%%\hline
%%%precedence(send colposcopy uncertain result, send biopsy sample) & 4 & 3 \\
%%%\hline
%%%\end{tabular}
%%%
%%%Model M2:\\
%%%
%%%\begin{tabular}{|l|l|l|}
%%%\hline
%%%Constraint & Traces & Paths \\
%%%\hline
%%%alternateresponse(send positive pap test result, take a colposcopy examination) & 4 & 1\\
%%%\hline
%%%chainprecedence(invite, take a pap test examination) & 3 & 2 \\
%%%\hline
%%%exclusivechoice(send pap test sample, reject) & 94 & 2 \\
%%%\hline
%%%succession(send colposcopy uncertain result, send biopsy sample) & 4 & 3\\
%%%\hline
%%%\end{tabular}

%%%A second example on this event log aims to demonstrate the important contribution of a negative preference, that is, a preference for the absence of a certain activity or template in the model. We applied this kind of preference on the “reject” action and we didn’t obtain any model satisfying the desired criterion. However, this result still offers a valuable information, as it guarantees that the “reject” activity is essential to distinguish between the positive and negative traces.

%%%In order to evaluate the performances of our approach we also report, for each preference of the previous examples, the time needed to retrieve different numbers of models.\\
%%%\begin{tabular}{|c|c|c|c|c|c|}
%%%\cline{3-6}
%%%\multicolumn{2}{c}{} & \multicolumn{4}{|c|}{Required Time (s)}\\
%%%\hline
%%%\multicolumn{2}{|c}{Log} & \multicolumn{2}{|c}{DREYERS} & \multicolumn{2}{|c|}{CERV}\\
%%%\hline
%%%\multicolumn{2}{|c|}{Preference} & architect activity & lawyer activity & precedence template & succession template\\
%%%\hline
%%%\multirow{5}*{Number of models} & 1 & 0.185 & 0.178 & 0.273 & 0.281\\
%%%\cline{2-6}
%%%& 5 & 0.285 & 0.439 & 0.646 & 0.666\\
%%%\cline{2-6}
%%%& 10 & 0.849 & 1.635 & 1.106 & 3.545\\
%%%\cline{2-6}
%%%& 20 & 2.938 & 3.517 & 3.066 & 18.159\\
%%%\cline{2-6}
%%%& 50 & 14.167 & 22.485 & 76.766 & 130.256\\
%%%\hline
%%%\end{tabular}
%%%
%%%This table shows that the execution time depends both on the log and the specific preference.



